\begin{itemize}
    \item \textbf{LCD Driver:}  
    \begin{itemize}
        \item Implemented in \texttt{lcd.h}, it includes functions such as:  
        \begin{itemize}
            \item \texttt{lcd\_init()}: Initializes the LCD.  
            \item \texttt{lcd\_clear()}: Clears the LCD display.  
            \item \texttt{lcd\_print()}: Prints a string or value on the LCD.  
        \end{itemize}
    \end{itemize}
    
    \item \textbf{Button Handling:}  
    \begin{itemize}
        \item Manages the detection of button presses, including numeric inputs (0-9) and special function keys.  
    \end{itemize}
    
    \item \textbf{Mathematical Evaluation:}  
    \begin{itemize}
        \item Utilizes a recursive descent parser for parsing arithmetic expressions.  
        \item Supports operations such as addition, subtraction, multiplication, and division.  
    \end{itemize}
    
    \item \textbf{Custom Mathematical Functions:}  
    \begin{itemize}
        \item Approximates trigonometric (\(\sin\), \(\cos\)) and logarithmic (\(\log\), \(\ln\)) functions using iterative methods.  
    \end{itemize}
\end{itemize}

\section*{Button Mapping and Modes}

Buttons facilitate numeric input and access various mathematical operations. The calculator operates in three modes:  

\begin{itemize}
    \item \textbf{Normal Mode:}  
    \begin{itemize}
        \item Allows entry of digits (0-9) and basic operations: \(+\), \(-\), \(*\), \(/\), \(=\), and \texttt{Backspace}.  
    \end{itemize}

    \item \textbf{Shift Mode:}  
    \begin{itemize}
        \item Provides access to advanced mathematical functions such as:  
        \begin{itemize}
            \item \(\sin\), \(\cos\), \(e^x\), and \(\sqrt{x}\).  
        \end{itemize}
    \end{itemize}

    \item \textbf{Extra Mode:}  
    \begin{itemize}
        \item Offers extended functionality with access to:  
        \begin{itemize}
            \item \(\arcsin\), \(\arccos\), \(\arctan\).  
            \item Logarithmic functions: \(\log\), \(\ln\).   
        \end{itemize}
    \end{itemize}
\end{itemize}
